AppPAL for Android

### 2015-06-02 paper.tex
Added to related work

% vim: set nowrap tw=0:
\documentclass[twoside,letterpaper]{soups}

\pdfpagewidth=8.5truein{}
\pdfpageheight=11truein{}

\usepackage{graphicx}
\renewcommand{\topfraction}{0.99} % be more aggressive about text around floats
\renewcommand{\floatpagefraction}{0.99}
\pagestyle{plain} % page numbers

@@

\title{AppPAL for Android}
\subtitle{Capturing and Checking App Installation Policies}
\numberofauthors{2}
\author{\alignauthor
  Joseph Hallett \\ \email{J.Hallett@sms.ed.ac.uk}
  \alignauthor
  David Aspinall \\ \email{David.Aspinal@ed.ac.uk}}

@@

\begin{figure}
  \centering
  \includegraphics[width=0.3\linewidth]{figures/facebook.png}
  \caption{Security information presented when a user downloads the Facebook app.}
\label{fig:facebook}
\end{figure}

@@

Using AppPAL we can make statements about apps:
\begin{lstlisting}
"Alice" says "com.rovio.angrybirds" isInstallable.
\end{lstlisting}
We can delegate decisions to others:
\begin{lstlisting}
"Alice" says "Bob" can-say 0 App isCool.
\end{lstlisting}.
We can run analysis tools, and solve constraints:
\begin{lstlisting}
"Claire" says App isMalware
  if App isAnApp
  where
    virusScanner(App) = True.
\end{lstlisting}

@@

Suppose \emph{Alice} works for her \emph{em\/}ployer \emph{Emma}.
Emma allows Alice to use her personal phone as a work phone, but she has some specific concerns.
\begin{itemize}
  \item Alice shouldn't have any apps that can track her movements.
    The testing labs are at a secret location and it mustn't be leaked.
  \item Apps should come from a reputable source, such as the Google Play Store.
  \item Emma uses an \ac{av} program by McAfee.
    It should check all apps before they're installed.
\end{itemize}

To ensure this policy is fully met Alice could promise to follow it.
Perhaps she might even sign a document promising to do so.
This is error-prone---what if she makes a mistake or misses something?
Emma's policy could be implemented using existing tools.
The enterprise tool \emph{Google's Device Policy for Android}\footnote{\url{https://play.google.com/store/apps/details?id=com.google.android.apps.enterprise.dmagent}}
  could configure Alice's device to disallow apps from outside the Google Play Store.
Various tools such as AppGuard~\citep{Backes:2012vm}, Dr.~Android~\&~Mr.~Hide~\citep{Jeon:2012ki} or AppFence~\citep{Hornyack:2011wq} can control the permissions or data an app can get.
These could be used to used to ensure no location data is ever obtained.
Alternately other tools like Kirin~\citep{Enck:2009ko}, Flowdroid~\citep{Fritz:2013vi} or DroidSafe~\citep{Gordon:2015et} could check that the locations are ever leaked to the web.
Various anti-virus programs are available for Android---one could be installed on Alice's phone checking against McAfee's signatures.

@@

Whilst we \emph{could} implement Emma's policy using existing tools, it isn't a great solution.
It isn't flexible, Alice needs to alter and remove the software on her phone if Emma changes her policy or she changes jobs.
It isn't clear what an app must do to be installed, or what checks have been done if it already has been put on the phone.
The link between Alice, the user; Emma, the policy setter; and the tools Emma trusts to implement her policy isn't immediately apparent.

Our research looks at the problem of picking software.
Given there are some apps you want and others you do not:
  how can you express your preferences in such a way that they can be enforced automatically?
How can we translate policy documents from natural language into something machine checkable?
Furthermore how can we show the trust relationships used to make these decisions clearly and precisely?

\section{Expressing Policies in AppPAL}
\label{sec:idea}

Returning to Alice and Emma's dilemma in \autoref{sec:problem}, we propose a solution.
Instead of using several different tools to enforce Emma's policy disjointedly, we use an authorization logic.
In \autoref{lst:corporate} we give an AppPAL policy implementing Emma's app concerns on Alice's phone.
A pictorial representation, is given in \autoref{fig:emmas_policy}; color is used to split statements by each speaker.

\begin{figure}
  \includegraphics{figures/bob-policy.eps}
  \caption{Pictorial representation of Alice's app install policy.}
\label{fig:emmas_policy}
\end{figure}

AppPAL is an instantiation of SecPAL~\citep{Becker:2006vh} for app installation policies.
SecPAL is a logic of authorization for access control decisions in distributed systems.
It has a clear and readable syntax, as well as rich mechanisms for delegation and constraints.
SecPAL has already been used as a basis for other policy languages in areas such as privacy preferences~\citep{Becker:2009ula} and data-sharing~\citep{Aziz:2011vt}.
We present AppPAL as a new version of SecPAL, targeting apps on mobile devices.

\begin{figure}
\begin{lstlisting}[numbers=left,
                   escapeinside={@}{@}]
"alice" says "emma" can-say inf
  App isInstallable. @\label{lst:corporate_1}@

"emma" says App isInstallable @\label{lst:corporate_2}@
  if "no-tracking-policy" isMetBy(App),
     "reputable-source-policy" isMetBy(App),
     "anti-virus-policy" isMetBy(App).

@@

"emma" says "anti-virus-policy" isMetBy(App) @\label{lst:corporate_5}@
  if App isAnApp
  where
    mcAfeeVirusCheck(App) = False.

@@

In \autoref{lst:corporate_1} Alice gives Emma the ability to specify whether an \code{App} (a variable) \code{isInstallable} (a predicate).
She allows her to delegate the decision further if she chooses (\code{can-say inf}).
Next in \autoref{lst:corporate_2} Emma specifies her concerns as policies to be met (the \code{isMetBy()} predicate that takes an app as its argument).
If Emma can be convinced all these policies are met then he will say the \code{App isInstallable}.
In \autoref{lst:corporate_3} and \autoref{lst:corporate_4} Emma specifies that an app meets the \code{reputable-source-policy} if the \code{App isReputable};
  with \code{"google-play"} specified as the decider of what is buyable or not.
This time Google is not allowed to delegate the decision further (\code{can-say 0}).
In other words Google is not allowed to specify Amazon as a supplier of apps as well.
Google must say what is buyable directly for Emma to accept it.
Emma specifies the \code{"anti-virus-policy"} in \autoref{lst:corporate_5}.
Here we use a constraint.
When checking the policy the \code{mcAfeeVirusCheck} should be run on the \code{App}.
Only if this returns \code{False} will the policy be met.
To specify the \code{"no-tracking-policy"} Emma says that the \code{"no-location-permissions"} rules implement the \code{"no-tracking-policy"} (\autoref{lst:corporate_6}).
Emma specifies this in \autoref{lst:corporate_7} by checking the app is missing two permissions.

Say Alice wants to install a new app (\code{com.facebook.katana}) on her phone.
To meet Emma's policy the AppPAL policy checker needs to collect statements to show the app meets the \code{isInstallable} predicate.
Specifically it needs:
\begin{itemize}
  \item\code{"emma" says "com.facebook.katana" isAnApp.}
    A simple typing statement that can be generated for all apps as they are encountered.
  \item\code{"google-play" says "com.facebook.katana" isReputable.}
    Required to convince Emma the app came from a reputable source.
    It should be able to obtain this statement from the Play store as the app is available there.
  \item\code{"emma" says "anti-virus-policy" isMetBy("com.facebook.katana").}
    She can obtain this by running the \ac{av} program on her app.
  \item\code{"emma" says "no-locations-permissions" isMetBy("com.facebook.katana").}
    Needed to show the App meets Emma's no-tracking-policy.
    Emma will say this if after examining the app the location permissions are missing.
\end{itemize}
These last two statements require the checker to do some extra checks to satisfy the constraints.
To get the third statement it must run the \ac{av} program on her app and check the result.
The results from the \ac{av} program may change with time as it's signatures are updated.
Consequently the checker must re-run this check every time it wants to obtain the statement connected to the constraint.
For the forth statement the checker needs to check the permissions of the app.
It does this by looking in the \texttt{MANIFEST.xml} inside the app itself.

@@

AppPAL is implemented as a library for Android.
A standalone Java version also exists for testing.
The parser is implemented using ANTLR4.
Code, tests, and prebuilt APKs are available from Github\footnote{\url{https://github.com/bogwonch/AppPAL}}.

@@

The structure of an AppPAL assertion is shown in \autoref{fig:assertion}.

\begin{figure}
  \centering
  \begin{equation*}\small
    \begin{array}{r l}
      \overbrace{\text{\ttfamily "user"}}^{\text{speaker}} &
      \text{\ttfamily says }\overbrace{\overbrace{\text{\ttfamily App }}^{\text{subject}}\overbrace{\text{\ttfamily isInstallable}}^{\text{predicate}}}^{\text{fact}} \\
      & \overbrace{\text{\ttfamily if App isFree}}^{\text{condition}} \\
      & \overbrace{\text{\ttfamily where hasPermission(App, "INTERNET") = True}}^{\text{constraint}}.
    \end{array}
  \end{equation*}
  \caption{Structure of an AppPAL assertion.}
\label{fig:assertion}
\end{figure}

@@

Specifically we define the following predicates:
\begin{description}
  \item[\texttt{App isInstallable}]
    Used to indicate that an App meets the install policy for the device.
    Showing an app satisfies this predicate is usually the goal of evaluating AppPAL.

@@

Splitting the decision as to whether an app is installable into a series of policies that must be met gives us flexibility in how the decision is made.
It allows us to describe multiple means of making the same decision, and provide backup routes when one means fails.
Some static analysis tools are not quick to run.
Even taking minutes to run an expensive (in terms of battery power) analysis can be undesirable.
If a user wants to download an app quickly they may not be willing to wait to check that a policy is met.

@@

"bob" says "no-locations-permissions" isMetBy(App)
  if App isAnApp
  where
    hasPermission(App, "ACCESS_FINE_LOCATION") = False,
    hasPermission(App, "ACCESS_COARSE_LOCATION") = False.

@@

"bob" says "location-taint-analysis" isMetBy(App)
  if App isAnApp
  where
    taintDroidCheckLeak(App, "Location", "Internet")
      = False.
\end{lstlisting}

Sometimes we might want to use location data.
For instance Bob might want to check that Alice is at her office.
Bob might use a location tracking app to do this.
Provided the app only talks to Bob, and it uses SSL correctly (something Mallodroid can check for~\citep{Fahl:2012dj}) he is happy to relax the policy.

\begin{lstlisting}
"bob" says "relaxed-no-tracking-policy" canActAs "no-tracking-policy".
"bob" says "relaxed-no-tracking-policy" isMetBy(App)
  if App hasCategory("tracking")
  where
    mallodroidSSLCheck(App) = False,
    connections(App) = "[https://bob.com]".
\end{lstlisting}

@@

From a user's perspective there was a delegation of trust to Lenovo to detect malware.
Lenovo then delegated further to McAfee to supply the antivirus checking; but left an exception that they would allow Superfish.
A user might wonder where the breach of trust occurred and who is to blame?
Is it the \ac{av} failing to spot the malware?
Has someone else configured something incorrectly?
How should they fix the problem?

@@

%
%We also add the following constraint functions which check simple properties of the app:
%In every case the check can be done by examining the app itself.
%\begin{description}
%  \item[\texttt{hasPermission(App, String)} : Bool]
%    Checks the permissions requested in the app's manifest to see if the string is included in them.
%
%  \item[\texttt{hasPackage(App, String)} : Bool]
%    Checks the app to see if a package called \emph{String} is included within it.
%    Examples might include aggressive advertisement frameworks a user wishes to ban.
%
%  \item[\texttt{hasNativeCode(App)} : Bool]
%    Checks to see whether the app contains native code.
%    Native (i.e. non-Dalvik) code is potentially more dangerous than the sandboxed Dalvik versions.
%    It is typically used for performance sensitive code (i.e. games).
%
%  \item[\texttt{isSignedBy(App, Speaker)} : Bool]
%    Checks the \texttt{CERT.RSA} used to sign the App and compares it with the certificate used to identify a speaker.
%
%  \item[\texttt{hasVersion(App)} : Int]
%    Gets the version number of an app.
%
%  \item[\texttt{released(App)} : Date]
%    Checks when the app was last updated and released.
%\end{description}
%
%We also add some constraints that do more complex checks:
%\begin{description}
%  \item[\texttt{googlePlayRating(App)} : Int]
%    Used to get the review score (out of 5) from the Google Play store.
%    This could be extended to other reviewers.
%
%  \item[tool\texttt{Check(App, Config)} : Bool]
%    Uses the \emph{tool} to run a check on an app.
%    This allows static analysis tools to run inferences on apps on demand.
%    Static analysis can be expensive for devices to perform.
%    An alternative would be to query a database of results~\citep{Noel:2014vt}.
%
%  \item[tool\texttt{CheckEvidence(App, Config, Evidence)} : Bool]
%    Digital evidence and proof~carrying~code~\citep{Necula:1996tr,Stark:2009uc} allow you to check an analysis result without repeating it.
%    In a mobile phone this can be preferable to doing the inference as draining batteries to perform checks is undesirable.
%    Tools, such as \emph{evicheck}~\citep{Noel:2014vt}, can produce digital evidence for Android apps.
%    This constraint function allows us to incorporate these tools into AppPAL policies.
%\end{description}

@@

To evaluate AppPAL we implement the SecPAL evaluation rules shown in \autoref{fig:rules}.
We do not use the DatalogC~\citep{Li:2003ix}~based translation and evaluation algorithm suggested by \citeauthor{Becker:2006vh}.
Rather we implement the rules directly in Java.
Brief pseudo-code is given in \autoref{fig:pseudocode}
Like \citeauthor{Becker:2006vh} we make use of an assertion context to store known statements.
We also store intermediate results to avoid re-computation.
On a mobile device memory is at a premium.
We would like to keep the context as small as possible.
To do this we allow some assertions to be imported at evaluation-time.

@@

In SecPAL there is no notion of type.
All entities are either constants (effectively strings) or variables.
There is no differentiation made between a \emph{voiced} constant (one who speaks),
and a \emph{subject} (one that is the subject of a verb-phrase, or predicate argument).
In AppPAL the subjects are usually apps or policies.
Apps do not utter facts.
Consequently there is no need to check them as potential speakers of \emph{can-say} statements.
Similarly if a constant never appears as a subject of a statement, as most speakers do not,
  there is no need to check it as a possible subject for a \emph{can-act-as} statement.
We pre-process the assertion context to find \emph{voiced} and \emph{subject} constants.

@@

In this section we give examples of policies and queries and show how they can be implemented in AppPAL.

\begin{description}
  \item[App white-listing by an IT department]
    Inside Emma's company employees in the IT department may white-list certain apps as installable.
    To do this Emma delegates to them to decide if something is installable.
    \begin{lstlisting}
"emma" says "it-department" can-say 0 App isInstallable.
    \end{lstlisting}
    Charlie and Diveena work in the IT department.
    When they have checked an app they say it is installable.
    \begin{lstlisting}
"emma" says "charlie" can-act-as "it-department".
"emma" says "diveena" can-act-as "it-department".

"charlie" says "com.facebook.katana" isInstallable.
"diveena" says "org.thoughtcrime.securesms" isInstallable.
    \end{lstlisting}

@@

    Alice's policy can be implemented by checking the permissions of the app, using constraints.
    \begin{lstlisting}
"alice" says App isInstallable
   if "permissionsPolicy" isMetBy(App).
"alice" says "permissionsPolicy" isMetBy(App)
   if App isAnApp
   where
      category(App, "Photography"),
      hasPermission(App, "ACCESS_COARSE_LOCATION") = False,
      hasPermission(App, "CAMERA") = True.
    \end{lstlisting}

@@

  \item[Digital Evidence]
    In \autoref{ssec:language} we described how digital evidence could be used to decide if policies are met.
    Suppose a digital evidence generating tool, such as \emph{Evicheck}, can be used to check a policy.
    For example that no audio can be recorded without the users consent~\citep{Seghir:2014uq}.
    Using this tool someone generates a proof certificate that this property is met.
    This could be the developer of the app, a third party certification service, or someone else;
      when we check the proof we will find out if it is valid or not.
    \begin{lstlisting}
"alice" says Anyone can-say 0
  "recording-consent" shownIsMetBy(Evidence, App).

@@

"alice" says "recording-consent" isMetBy(App)
  if "recording-consent" shownIsMetBy(Evidence, App)
  where
    evicheckCheckEvidence(Evidence, App) = True.
    \end{lstlisting}
    In the example here Alice allows anyone to present her with evidence for her policy.
    A \ac{skb} offers an some evidence, and Alice accepts it as if it checks out.
\end{description}

\section{Related work \comment{1--2 pages}}

We picked SecPAL as the basis for AppPAL because it was readable, extensible, and seemed to be a good fit for the problem we were trying to solve~\citep{Hallett:2014un}.
Further more attempts to apply it to other areas had already succeeded~\citep{Aziz:2011vt,Humphrey:2007wc}.
Other policy languages such as RT~\citep{Li:2003ua}, or Cassandra~\citep{Becker:2004il} might also have made a good starting point.
However neither had the precise set of features we we're looking for (namely explicit speakers and external checking functions).
Other work on SecPAL has added various features such as existential quantification~\citep{Becker:2009vt} and ultimately becoming the DKAL family of policy languages~\citep{Gurevich:2008fz,Gurevich:Qo5E3M3}.
\citeauthor{Gurevich:2008fz} showed that SecPAL was a subset of DKAL (minus the \emph{can-act-as} statement).
DKAL also contains more modalities than \emph{says}.

>>>>

% vim: set nowrap tw=0:
\documentclass[]{scrartcl}

% \pdfpagewidth=8.5truein{}
% \pdfpageheight=11truein{}

\usepackage{graphicx}
% \renewcommand{\topfraction}{0.99} % be more aggressive about text around floats
% \renewcommand{\floatpagefraction}{0.99}
% \pagestyle{plain} % page numbers

@@

\title{AppPAL for Android}
\subtitle{Capturing and Checking App Installation Policies}
% \numberofauthors{2}
\author{Joseph Hallett \and David Aspinall }

@@

% \begin{figure}
%   \centering
%   \includegraphics[width=0.3\linewidth]{figures/facebook.png}
%   \caption{Security information presented when a user downloads the Facebook app.}
% \label{fig:facebook}
% \end{figure}

@@

Using AppPAL we can make statements about apps:
\begin{lstlisting}
"Alice" says "com.rovio.angrybirds" isRunnable.
\end{lstlisting}
We can delegate decisions to others:
\begin{lstlisting}
"Alice" says "Bob" can-say 0 App isCool.
\end{lstlisting}
We can run analysis tools, and solve constraints:
\begin{lstlisting}
"Claire" says App isMalware
  if App isAnApp
  where
    virusScanner(App) = true.
\end{lstlisting}

@@

Suppose \emph{Alice} works for her \emph{em\/}ployer \emph{Emma}.
Emma allows Alice to use her personal phone as a work phone, but she has some specific concerns.
\begin{itemize}
  \item Alice shouldn't run any apps that can track her movements.
    The testing labs are at a secret location and it mustn't be leaked.
  \item Apps should come from a reputable source, such as the Google Play Store.
  \item Emma uses an \ac{av} program by McAfee.
    It should check all apps before they're installed.
\end{itemize}

To ensure this policy is fully met Alice could promise to follow it.
Perhaps she might even sign a document promising to do so.
This is error-prone---what if she makes a mistake or misses an app that breaks her policy?
Emma's policy could be implemented using existing tools.
The enterprise tool \emph{Google's Device Policy for Android}\footnote{\url{https://play.google.com/store/apps/details?id=com.google.android.apps.enterprise.dmagent}}
  could configure Alice's device to disallow apps from outside the Google Play Store.
Various tools such as AppGuard~\citep{Backes:2012vm}, Dr.~Android~\&~Mr.~Hide~\citep{Jeon:2012ki} or AppFence~\citep{Hornyack:2011wq} can control the permissions or data an app can get.
These could be used to used to ensure no location data is ever obtained.
Alternately other tools like Kirin~\citep{Enck:2009ko}, Flowdroid~\citep{Fritz:2013vi} or DroidSafe~\citep{Gordon:2015et} could check that the locations are ever leaked to the web.
Various anti-virus programs are available for Android---one could be installed on Alice's phone checking against McAfee's signatures.

@@

Whilst we \emph{could} implement Emma's policy using existing tools, it isn't a great solution.
It isn't flexible, Alice needs to alter and remove the software on her phone if Emma changes her policy or she changes jobs.
It isn't clear what an app must do to be run, or what checks have been done if it already has been put on the phone.
The link between Alice, the user; Emma, the policy setter; and the tools Emma trusts to implement her policy isn't immediately apparent.

What happens when Alice goes home?
Emma shouldn't be able to overly control what Alice does in her private life.
Alice might not be allowed to use location tracking apps at work, but at home she might want to (to meet friends, track jogging routes or find restaurants for example).
The some mobile OSs allow app permissions to be enabled and disable at run time.
Can we enforce different policies at different times?

Our research looks at the problem of picking software.
Given there are some apps you want and others you do not:
  how can you express your preferences in such a way that they can be enforced automatically?
How can we translate policy documents from natural language into a machine checkable form?
Furthermore how can we show the trust relationships used to make these decisions clearly and precisely?

\section{Expressing policies in AppPAL}
\label{sec:idea}

Returning to Alice and Emma's dilemma in \autoref{sec:problem}, we propose a solution.
Instead of using several different tools to enforce Emma's policy disjointedly, we use an authorization logic.
In \autoref{lst:corporate} we give an AppPAL policy implementing Emma's app concerns on Alice's phone.
% A pictorial representation, is given in \autoref{fig:emmas_policy}; color splits statements by each speaker.

% \begin{figure}
%   \includegraphics{figures/bob-policy.eps}
%   \caption{Pictorial representation of Alice's app install policy.}
% \label{fig:emmas_policy}
% \end{figure}

AppPAL is an instantiation of SecPAL~\citep{Becker:2006vh} for app running policies.
SecPAL is a logic of authorization for access control decisions in distributed systems.
It has a clear and readable syntax, as well as rich mechanisms for delegation and constraints.
SecPAL has already been used as a basis for other policy languages in areas such as privacy preferences~\citep{Becker:2009ula} and data-sharing~\citep{Aziz:2011vt}.
We present AppPAL as a new version of SecPAL, targeting apps on mobile devices.

\begin{figure}
\begin{lstlisting}[numbers=left,
                   escapeinside={@}{@}]
"alice" says "emma" can-say inf
  App isRunnable. @\label{lst:corporate_1}@

"emma" says App isRunnable @\label{lst:corporate_2}@
  if "no-tracking-policy" isMetBy(App),
     "reputable-source-policy" isMetBy(App),
     "anti-virus-policy" isMetBy(App).

@@

"emma" says "anti-virus-policy" isMetBy(App) @\label{lst:corporate_5}@
  if App isAnApp
  where
    mcAfeeVirusCheck(App) = false.

@@

In \autoref{lst:corporate_1} Alice gives Emma the ability to specify whether an \code{App} (a variable) \code{isRunnable} (a predicate).
She allows her to delegate the decision further if she chooses (\code{can-say inf}).
Next in \autoref{lst:corporate_2} Emma specifies her concerns as policies to be met (the \code{isMetBy()} predicate that takes an app as its argument).
If Emma can be convinced all these policies are met then he will say the \code{App isRunnable}.
In \autoref{lst:corporate_3} and \autoref{lst:corporate_4} Emma specifies that an app meets the \code{reputable-source-policy} if the \code{App isReputable};
  with \code{"google-play"} specified as the decider of what is buyable or not.
This time Google is not allowed to delegate the decision further (\code{can-say 0}).
In other words Google is not allowed to specify Amazon as a supplier of apps as well.
Google must say what is buyable directly for Emma to accept it.
Emma specifies the \code{"anti-virus-policy"} in \autoref{lst:corporate_5}.
Here we use a constraint.
When checking the policy the \code{mcAfeeVirusCheck} should be run on the \code{App}.
Only if this returns \code{false} will the policy be met.
To specify the \code{"no-tracking-policy"} Emma says that the \code{"no-location-permissions"} rules implement the \code{"no-tracking-policy"} (\autoref{lst:corporate_6}).
Emma specifies this in \autoref{lst:corporate_7} by checking the app is missing two permissions.

Say Alice wants to install a new app (\code{com.facebook.katana}) on her phone.
To meet Emma's policy the AppPAL policy checker needs to collect statements to show the app meets the \code{isRunnable} predicate.
Specifically it needs:
\begin{itemize}
  \item\code{"emma" says "com.facebook.katana" isAnApp.}
    A simple typing statement that can be generated for all apps as they are encountered.
  \item\code{"google-play" says "com.facebook.katana" isReputable.}
    Required to convince Emma the app came from a reputable source.
    It should be able to obtain this statement from the Play store as the app is available there.
  \item\code{"emma" says "anti-virus-policy" isMetBy("com.facebook.katana").}
    She can obtain this by running the \ac{av} program on her app.
  \item\code{"emma" says "no-locations-permissions" isMetBy("com.facebook.katana").}
    Needed to show the App meets Emma's no-tracking-policy.
    Emma will say this if after examining the app the location permissions are missing.
\end{itemize}
These last two statements require the checker to do some extra checks to satisfy the constraints.
To get the third statement it must run the \ac{av} program on her app and check the result.
The results from the \ac{av} program may change with time as it's signatures are updated.
Consequently the checker must re-run this check every time it wants to obtain the statement connected to the constraint.
For the forth statement the checker needs to check the permissions of the app.
It does this by looking in the \texttt{MANIFEST.xml} inside the app itself.

@@

AppPAL is implemented as a library for Android and Java.
The parser is implemented using ANTLR4.
Code and build instructions are available from Github\footnote{\url{https://github.com/bogwonch/libAppPAL}}.

@@

The structure of an AppPAL assertion can be seen in \autoref{fig:assertion}.

\begin{figure}
  \centering
  \begin{equation*}\small
    \begin{array}{r l}
      \overbrace{\text{\ttfamily "user"}}^{\text{speaker}} &
      \text{\ttfamily says }\overbrace{\overbrace{\text{\ttfamily App }}^{\text{subject}}\overbrace{\text{\ttfamily isRunnable}}^{\text{predicate}}}^{\text{fact}} \\
      & \overbrace{\text{\ttfamily if App isFree}}^{\text{condition}} \\
      & \overbrace{\text{\ttfamily where hasPermission(App, "INTERNET") = true}}^{\text{constraint}}.
    \end{array}
  \end{equation*}
  \caption{Structure of an AppPAL assertion.}
\label{fig:assertion}
\end{figure}

@@

Specifically we define the following predicates:
\begin{description}
  \item[\texttt{App isRunnable}]
    Used to indicate that an App meets the install policy for the device.
    Showing an app satisfies this predicate is usually the goal of evaluating AppPAL.

@@

Splitting the decision about whether an app is runnable into a series of policies that must be met gives us flexibility in how the decision is made.
It allows us to describe multiple means of making the same decision, and provide backup routes when one means fails.
Some static analysis tools are not quick to run.
Even taking minutes to run a battery draining analysis can be undesirable.
If a user wants to download an app quickly they may not be willing to wait to check that a policy is met.

@@

"bob" says "no-locations-permissions" isMetBy(App)
  if App isAnApp
  where
    hasPermission(App, "ACCESS_FINE_LOCATION") = false,
    hasPermission(App, "ACCESS_COARSE_LOCATION") = false.

@@

"bob" says "location-taint-analysis" isMetBy(App)
  if App isAnApp
  where
    taintDroidCheckLeak(App, "Location", "Internet")
      = false.
\end{lstlisting}

Sometimes we might want to use location data.
For instance Bob might want to check that Alice is at her office.
Bob might track Alice using a location tracking app.
Provided the app only talks to Bob, and it uses SSL correctly (which Mallodroid can check for~\citep{Fahl:2012dj}) he is happy to relax the policy.

\begin{lstlisting}
"bob" says "relaxed-no-tracking-policy" canActAs "no-tracking-policy".
"bob" says "relaxed-no-tracking-policy" isMetBy(App)
  if App hasCategory("tracking")
  where
    mallodroidSSLCheck(App) = false,
    connections(App) = "[https://bob.com]".
\end{lstlisting}

@@

From a user's perspective there was a delegation of trust to Lenovo to detect malware.
Lenovo then delegated further to McAfee to supply the antivirus checking; but left an exception that they would allow Superfish.
A user might wonder where the breach of trust occurred and who is to blame?
Is it the \ac{av} failing to spot the malware?
Has someone else configured the software incorrectly?
How should they fix the problem?

@@

@@

To evaluate AppPAL we implement the SecPAL evaluation rules shown in \autoref{fig:rules}.
We do not use the DatalogC~\citep{Li:2003ix}~based translation and evaluation algorithm suggested by \citeauthor{Becker:2006vh}.
Rather we implement the rules directly in Java.
Brief pseudo-code is given in \autoref{fig:pseudocode}
Like \citeauthor{Becker:2006vh} we make use of an assertion context to store known statements.
We also store intermediate results to avoid re-computation.
On a mobile device memory is at a premium.
We would like to keep the context as small as possible.
For some assertions (like \code{isAnApp}) we derive them by checking the arguments at evaluation time.

@@

In SecPAL there is no type.
All entities are either constants (effectively strings) or variables.
There is no differentiation made between a \emph{voiced} constant (who says assertions),
and a \emph{subject} (the subject of a verb-phrase, or predicate argument).
In AppPAL the subjects are usually apps or policies.
Apps do not utter facts.
Consequently there is no need to check them as potential speakers of \emph{can-say} statements.
Similarly if a constant never appears as a subject of a statement, as most speakers do not,
  there is no need to check it as a possible subject for a \emph{can-act-as} statement.
We pre-process the assertion context to find \emph{voiced} and \emph{subject} constants.

@@

We give examples of policies and queries and show how they can be implemented in AppPAL.

\begin{description}
  \item[Policies for home and work]
    In \autoref{sec:introduction} we said Alice might like to have separate rules for home and work.
    Emma may insist that her policies come into effect whenever Alice is at work, but is okay with Alice breaking them after 5~pm or when she isn't at work.
    To implement this we add rules that use constraints to decide whether an app should be run.
    Alice already has the Emma's work policy as the default.
    Now she just needs to add an exception to the rule.
    \begin{lstlisting}
"alice" says App isRunnable
  if "home-policy" isMetBy(App)
  where TimeOfDay() > 17.00.

"alice" says App isRunnable
  if "home-policy" isMetBy(App)
  where At("work") = false.
    \end{lstlisting}

  \item[App white-listing by an IT department]
    Inside Emma's company employees in the IT department may white-list certain apps as runnable.
    Emma delegates to them to decide if an app is runnable.
    \begin{lstlisting}
"emma" says "it-department" can-say 0 App isRunnable.
    \end{lstlisting}
    Charlie and Diveena work in the IT department.
    When they have checked an app they say it is runnable.
    \begin{lstlisting}
"emma" says "charlie" can-act-as "it-department".
"emma" says "diveena" can-act-as "it-department".

"charlie" says "com.facebook.katana" isRunnable.
"diveena" says "org.thoughtcrime.securesms" isRunnable.
    \end{lstlisting}

@@

    Alice's policy can be implemented by checking the permissions of the app, using constraints.
    \begin{lstlisting}
"alice" says App isRunnable
   if "permissionsPolicy" isMetBy(App).
"alice" says "permissionsPolicy" isMetBy(App)
   if App isAnApp
   where
      category(App, "Photography"),
      hasPermission(App, "ACCESS_COARSE_LOCATION") = false,
      hasPermission(App, "CAMERA") = true.
    \end{lstlisting}

@@

  \item[Digital Evidence]
    In \autoref{ssec:language} we described how digital evidence could be used to decide if policies are met.
    Suppose a digital evidence generating tool, such as \emph{Evicheck}, can be used to check a policy.
    For example that no audio can be recorded without the users consent~\citep{Seghir:2014uq}.
    Using this tool someone generates a proof certificate that this property is met.
    This might be the developer of the app, a third party certification service, or someone else;
      when we check the proof we will discover if it is valid or not.
    \begin{lstlisting}
"alice" says Anyone can-say 0
  "recording-consent" shownIsMetBy(Evidence, App).

@@

"alice" says "recording-consent" isMetBy(App)
  if "recording-consent" shownIsMetBy(Evidence, App)
  where
    evicheckCheckEvidence(Evidence, App) = true.
    \end{lstlisting}
    In the example here Alice allows anyone to present her with evidence for her policy.
    A \ac{skb} offers some evidence, and Alice accepts it as if it checks out.
\end{description}

\section{Related work}

Authorization logics have been successfuly used to enforce policies in several other domains.
The earliest such logic, PolicyMaker~\cite{Blaze:dj}, was general, if undecidable.
Logics that followed like KeyNote~\citep{Blaze:1999fa} and SPKI/SDSI~\citep{Ellison:1999ui} looked at public key infrastructure.
The RT-languages~\citep{Li:2002if,Li:2003ix,Li:2003to} were designed for credential management.
Cassandra~\citep{Becker:2004fi} was used to model trust relationships in the british national health service.

SELinux is used to describe policies for Linux processes, and for access control (on top of the Linux discretionary controls).
It was ported to Android~\citep{Smalley:2013vl} and is used in the implementation of the permissions system.
Google also offer an app for configuring a \emph{device policy}\footnote{\url{https://play.google.com/store/apps/details?id=com.google.android.apps.enterprise.dmagent}}.
This lets businesses to configure company owned devices to be trackable, remote lockable, set passwords and sync with their servers.
It cannot be used to describe policies about apps, or describe trust relationships, however.

The SecPAL language was initially used for access control in distributed systems.
We picked SecPAL as the basis for AppPAL because it was readable, extensible, and seemed to be a good fit for the problem we were trying to solve~\citep{Hallett:2014un}.
It has already been used to describe data usage policies~\citep{Aziz:2011vt} and inside Grid data systems~\cite{Humphrey:2007wc}.
Other work on SecPAL has added various features such as existential quantification~\citep{Becker:2009vt} and ultimately becoming the DKAL family of policy languages~\citep{Gurevich:2008fz,Gurevich:Qo5E3M3}.
\citeauthor{Gurevich:2008fz} showed that SecPAL was a subset of DKAL (minus the \emph{can-act-as} statement).
DKAL also contains more modalities than \emph{says}.

One tool, Kirin~\citep{Enck:2009ko}, also created a policy language and tool for enforcing app installation policies.
Kirin's policies were concerned with preventing malware.
Policy authors could specify combinations of permissions that should not appear together.
For example an author might wish to stop malware sending premium rate text messages.
To might implement this by restricting an app having both the \texttt{SEND\_SMS} and \texttt{WRITE\_SMS} permissions (\autoref{fig:kirin}).
Using this approach they found vulnerabilities in Android, but were ultimately limited by being restricted to permissions and broadcast events.
\begin{figure}
\begin{lstlisting}
restrict permission [SEND_SMS] and permission [WRITE_SMS]
\end{lstlisting}
\begin{lstlisting}
"user" says "no-write-send-sms" isMetBy(App)
  where hasPermission(App, "SEND_SMS") = false.
"user" says "no-write-send-sms" isMetBy(App)
  where hasPermission(App, "WRITE_SMS") = false.
\end{lstlisting}
\caption{Kirin and AppPAL policies for stopping apps monetized by premium rate text messages.}
\label{fig:kirin}
\end{figure}

Work by \citeauthor{Sadeh:2014vq} looked at modelling users' willingness to grant different kinds of permissions~\citep{Sadeh:2014vq}.
They found four clusters of different users each with different kinds of policies:
  conservatives who do not like granting many permissions,
  advanced users who seem to have more complex policies,
  unconcerned users who'll comfortably grant anything,
  and fence-sitters (about half of all users) who seemed not to have varying opinions.
This suggests that users are applying policies and do care how apps behave.
\citeauthor{Sadeh:2014vq} speculate that the fence-sitters seemed not to care because of \emph{warning fatigue}.
This suggests a system like AppPAL that can reduce user interaction with permission warnings might be interesting.
Using a policy unconcerned and fence-sitting users could take up more privacy preserving policies without having to change their behavior.
